---
title: "European Cars"
output: html_document
date: "2025-02-13"
---

```{r setup, include=FALSE}
#install.packages("psych")
#install.packages("corrplot")
library(corrplot)
library(psych)
library(tidyverse)
library(skimr)
library(dplyr)
```

## R Markdown

```{r}
EUcars <- read_csv("car_prices.csv", na = c("", NA, "N/A"))
EUcars = EUcars |> drop_na() |> filter (Brand == "Fiat", Gearbox != "Semi-automatic", Drivetrain != "Rear") |> select(Model,Price,Kilometers,Power, Seats,Doors,Cylinders,Gearbox,Year,Drivetrain,Country)

EUcars <- replace(EUcars, EUcars == "Automatic", "0")
EUcars <- replace(EUcars, EUcars == "Manual", "1")
EUcars <- replace(EUcars, EUcars == "4WD", "1")
EUcars <- replace(EUcars, EUcars == "Front", "0")
EUcars <- replace(EUcars, EUcars == "AT", "1")
EUcars <- replace(EUcars, EUcars == "BE", "2")
EUcars <- replace(EUcars, EUcars == "DE", "3")
EUcars <- replace(EUcars, EUcars == "IT", "4")
EUcars <- replace(EUcars, EUcars == "NL", "5")

EUcars$Gearbox = as.numeric(EUcars$Gearbox)
EUcars$Drivetrain = as.numeric(EUcars$Drivetrain)
EUcars$Country = as.numeric(EUcars$Country)

```

```{r}
stripchart(EUcars[,2], method = "stack", pch = 16, col = "blue",xlab="Price")
stripchart(EUcars[,3], method = "stack", pch = 16, col = "blue",xlab="Kilometers")
stripchart(EUcars[,4], method = "stack", pch = 16, col = "blue",xlab="Power")
stripchart(EUcars[,5], method = "stack", pch = 16, col = "blue",xlab="Seats")
stripchart(EUcars[,6], method = "stack", pch = 16, col = "blue",xlab="Doors")
stripchart(EUcars[,7], method = "stack", pch = 16, col = "blue",xlab="Cylinders")
stripchart(EUcars[,8], method = "stack", pch = 16, col = "blue",xlab="Gearbox")
stripchart(EUcars[,9], method = "stack", pch = 16, col = "blue",xlab="Year")
stripchart(EUcars[,10], method = "stack", pch = 16, col = "blue",xlab="Drivetrain")
stripchart(EUcars[,11], method = "stack", pch = 16, col = "blue",xlab="Country")


describe(EUcars)
```

## Visualize and summarize the covariability between each pair of dimensions

Our data has some variables that are highly correlated. Outstanding correlation includes: Kilometers and Price with a strong negative correlation, Price and Power with a strong positive correlation, Price and Year with a strong positive correlation, Kilometers and Year with a strong negative correlation and Power and Seats with a strong positive correlation. Some variables such as Cylinders and Price as well as Drivetrain and Price have very little correlation.All the other varibles have corelation somewhere in between.  

```{r}
cov_matrix<-cov(EUcars[,2:11])
cov_matrix
pairs(cov(EUcars[,2:11]), col = "blue", pch = 16,main="Covariability Scatter Plot")

cor_matrix<-cor(EUcars[,2:11])
cor_matrix
pairs(cor(EUcars[,2:11]), col = "blue", pch = 16,main="Correlation Scatter Plot")
corrplot(cor_matrix, title="Correlation Plot")
```

## Identify potential two-dimensional outliers and investigate their influence

```{r}
m <- mahalanobis(x = EUcars[,2:3],
center = c(10597.53, 81334.99),
cov = cov(EUcars[,2:3]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,2:3])
cor(EUcars[c(1:58, 60:117, 119:121, 123:146, 148:224, 226:242),2:3])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,4)],
center = c(10597.53, 92.29),
cov = cov(EUcars[,c(2,4)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,4)])
cor(EUcars[c(2:55, 57:65, 67:146, 148:158, 160:161, 163:167, 169:224, 226:242),c(2,4)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,5)],
center = c(10597.53, 4.56),
cov = cov(EUcars[,c(2,5)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,5)])
cor(EUcars[c(2:42, 44:89, 91:103, 105:119, 121, 123:146, 148:158, 160:206, 208:224, 226:242),c(2,5)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,6)],
center = c(10597.53, 4.36),
cov = cov(EUcars[,c(2,6)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,6)])
cor(EUcars[c(1:146, 148:161, 163:224, 226:242),c(2,6)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,7)],
center = c(10597.53, 3.60),
cov = cov(EUcars[,c(2,7)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,7)])
cor(EUcars[c(1:57, 59:146, 148:161, 163:224, 226:241),c(2,7)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,8)],
center = c(10597.53, 0.86),
cov = cov(EUcars[,c(2,8)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,8)])
cor(EUcars[c(2:19, 21:41, 43:46, 48:112, 114:146, 148:161, 163:224, 226:242),c(2,8)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,9)],
center = c(10597.53, 2017.32),
cov = cov(EUcars[,c(2,9)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,9)])
cor(EUcars[c(1:27, 29:112, 114:138, 140:142, 144:146, 148:194, 196:213, 215:224, 226:242),c(2,9)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,10)],
center = c(10597.53, 0.04),
cov = cov(EUcars[,c(2,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:55, 57:116, 118:146, 148:152, 154:158, 160:164, 166:167, 169:224, 226:242),c(2,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(2,11)],
center = c(10597.53, 3.86),
cov = cov(EUcars[,c(2,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(2,11)])
cor(EUcars[c(1:7, 9:23, 25:95, 97:105, 107:112, 114:146, 148:174, 176:224, 226:242),c(2,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,4)],
center = c(81334.99, 92.29),
cov = cov(EUcars[,c(3,4)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,4)])
cor(EUcars[c(2:65, 67:117, 119:167, 169:242),c(3,4)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,5)],
center = c(81334.99, 4.56),
cov = cov(EUcars[,c(3,5)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,5)])
cor(EUcars[c(2:42, 44:89, 91:103, 105:117, 119, 121, 123:146, 148:167, 169:206, 208:242),c(3,5)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,6)],
center = c(81334.99, 4.36),
cov = cov(EUcars[,c(3,6)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,6)])
cor(EUcars[c(1:117, 119:242),c(3,6)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,7)],
center = c(81334.99, 3.60),
cov = cov(EUcars[,c(3,7)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,7)])
cor(EUcars[c(1:51, 53:57, 59:117, 119:161, 163:241),c(3,7)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,8)],
center = c(81334.99, 0.86),
cov = cov(EUcars[,c(3,8)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,8)])
cor(EUcars[c(2:117, 119:167, 169:242),c(3,8)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,9)],
center = c(81334.99, 2017.32),
cov = cov(EUcars[,c(3,9)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,9)])
cor(EUcars[c(1:27, 29:58, 60:117, 119:134, 136:138, 140:142, 144:194, 196:213, 215:242),c(3,9)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,10)],
center = c(81334.99, 0.04),
cov = cov(EUcars[,c(3,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:55, 57:116, 119:152, 154:158, 160:164, 166, 169:242),c(3,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(3,11)],
center = c(81334.99, 3.86),
cov = cov(EUcars[,c(3,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(3,11)])
cor(EUcars[c(1:23, 25:117, 119:174, 176:242),c(3,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(4,5)],
center = c(92.29, 4.56),
cov = cov(EUcars[,c(4,5)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(4,5)])
cor(EUcars[c(2:18, 20:42, 44:55, 57:65, 67:89, 91:103, 105:119, 121, 123:146, 148:158, 160:167, 169:206, 208:242),c(4,5)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(4,6)],
center = c(92.29, 4.36),
cov = cov(EUcars[,c(4,6)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(4,6)])
cor(EUcars[c(1:65, 67:158, 160:242),c(4,6)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(4,7)],
center = c(92.29, 3.60),
cov = cov(EUcars[,c(4,7)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(4,7)])
cor(EUcars[c(1:57, 59:161, 163:241),c(4,7)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(4,8)],
center = c(92.29, 0.86),
cov = cov(EUcars[,c(4,8)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(4,8)])
cor(EUcars[c(2:15, 17:18, 21:55, 57:65, 67:80, 82:84, 86:133, 135:151, 153:158, 160:161, 163:167, 169:197, 199:237, 239:241),c(4,8)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(4,9)],
center = c(92.29, 2017.32),
cov = cov(EUcars[,c(4,9)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(4,9)])
cor(EUcars[c(2:18, 20:27, 29:65, 67:138, 140:142, 144:158, 160:167, 169:213, 215:242),c(4,9)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(4,10)],
center = c(92.29, 0.04),
cov = cov(EUcars[,c(4,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(4,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:55, 57:65, 67:116, 118:152, 154:158, 160:164, 166:167, 169:242),c(4,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(4,11)],
center = c(92.29, 3.86),
cov = cov(EUcars[,c(4,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(4,11)])
cor(EUcars[c(1:23, 25:65, 67:80, 82:99, 101:105, 107:174, 176:242),c(4,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(5,6)],
center = c(4.56, 4.36),
cov = cov(EUcars[,c(5,6)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(5,6)])
cor(EUcars[c(1:42, 44:89, 91:103, 105:119, 121, 123:146, 148:158, 160:206, 208:242),c(5,6)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(5,7)],
center = c(4.56, 3.60),
cov = cov(EUcars[,c(5,7)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(5,7)])
cor(EUcars[c(1:42, 44:57, 59:89, 91:103, 105:119, 121, 123:146, 148:161, 163:206, 208:241),c(5,7)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(5,8)],
center = c(4.56, 0.86),
cov = cov(EUcars[,c(5,8)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(5,8)])
cor(EUcars[c(2:18, 20:42, 44:55, 57:89, 91:103, 105:119, 121, 123:146, 149:158, 160:167, 169:206, 208:242),c(5,8)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(5,9)],
center = c(4.56, 2017.32),
cov = cov(EUcars[,c(5,9)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(5,9)])
cor(EUcars[c(1:18, 20:27, 29:42, 44:89, 91:103, 105:119, 121, 123:138, 140:142, 144:146, 149:158, 160:167, 169:206, 208:213, 215:242),c(5,9)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(5,10)],
center = c(4.56, 0.04),
cov = cov(EUcars[,c(5,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(5,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:42, 44:55, 57:89, 91:103, 105:116, 118, 119, 121, 123:146, 149:152, 154:158, 160:164, 166, 169:206, 208:242),c(5,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(5,11)],
center = c(4.56, 3.86),
cov = cov(EUcars[,c(5,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(5,11)])
cor(EUcars[c(1:18, 20:42, 44:89, 91:103, 105:119, 121, 123:146, 148:158, 160:174, 176:206, 208:242),c(5,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(6,7)],
center = c(4.36, 3.60),
cov = cov(EUcars[,c(6,7)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(6,7)])
cor(EUcars[c(1:33, 35:57, 59:161, 163:241),c(6,7)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(6,8)],
center = c(4.36, 0.86),
cov = cov(EUcars[,c(6,8)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(6,8)])
cor(EUcars[c(1:84, 86:161, 163:242),c(6,8)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(6,9)],
center = c(4.36, 2017.32),
cov = cov(EUcars[,c(6,9)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(6,9)])
cor(EUcars[c(1:27, 29:138, 140:142, 144:161, 163:213, 215:242),c(6,9)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(6,10)],
center = c(4.36, 0.04),
cov = cov(EUcars[,c(6,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(6,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:55, 57:116, 118:152, 154:158, 160:164, 166, 169:242),c(6,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(6,11)],
center = c(4.36, 3.86),
cov = cov(EUcars[,c(6,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(6,11)])
cor(EUcars[c(1:23, 25:33, 35:80, 82:105, 107:174, 176:242),c(6,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(7,8)],
center = c(3.60, 0.86),
cov = cov(EUcars[,c(7,8)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(7,8)])
cor(EUcars[c(1:57, 59:161, 163:241),c(7,8)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(7,9)],
center = c(3.60, 2017.32),
cov = cov(EUcars[,c(7,9)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(7,9)])
cor(EUcars[c(1:27, 29:57, 59:138, 140:142, 144:161, 163:213, 215:241),c(7,9)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(7,10)],
center = c(3.60, 0.04),
cov = cov(EUcars[,c(7,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(7,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:55, 57, 59:116, 118:152, 154:158, 160, 161, 163:164, 166, 167, 169:241),c(7,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(7,11)],
center = c(3.60, 3.86),
cov = cov(EUcars[,c(7,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(7,11)])
cor(EUcars[c(1:57, 59:161, 163:174, 176:241),c(7,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(8,9)],
center = c(0.86, 2017.32),
cov = cov(EUcars[,c(8,9)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(8,9)])
cor(EUcars[c(1:19, 21:27, 29:138, 140:142, 144:213, 215:242),c(8,9)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(8,10)],
center = c(0.86, 0.04),
cov = cov(EUcars[,c(8,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(8,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:55, 57:116, 118:152, 154:158, 160, 161, 163:164, 166, 167, 169:242),c(8,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(8,11)],
center = c(0.86, 3.86),
cov = cov(EUcars[,c(8,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(8,11)])
cor(EUcars[c(1:19, 21:80, 82:174, 176:242),c(8,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(9,10)],
center = c(2017.32, 0.04),
cov = cov(EUcars[,c(9,10)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(9,10)])
cor(EUcars[c(2:16, 18, 20:24, 26:27, 29:55, 57:116, 118:138, 140:142, 144:152, 154:158, 160, 161, 163:164, 166, 167, 169:213, 215:242),c(9,10)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(9,11)],
center = c(2017.32, 3.86),
cov = cov(EUcars[,c(9,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(9,11)])
cor(EUcars[c(1:23, 25, 26:38, 40:138, 140:142, 144:174, 176:213, 215:242),c(9,11)])
```


```{r}
m <- mahalanobis(x = EUcars[,c(10,11)],
center = c(0.04, 3.86),
cov = cov(EUcars[,c(10,11)]))
which(m > qchisq(p = 1 - 0.01, df = 2))
cor(EUcars[,c(10,11)])
cor(EUcars[c(2:16, 18, 20:24, 26:55, 57:116, 118:152, 154:158, 160, 161, 163:164, 166, 167, 169:174, 176:242),c(10,11)])
```

In summary the only significant differences at the 1% level are with Price and Drivetrain, Kilometers with doors and drivetrain, Power with drivetrain, seats with year and drivetrain, Doors with cylinders, gearbox, drivetrain, and country, Cylinders with drivetrain, Gearbox with drivetrain, Year and drivetrain, and country and drivetrain. None of these are our highly correlated variables though.
