---
title: "European Cars"
output: html_document
date: "2025-02-13"
---

```{r setup, include=FALSE}
#install.packages("psych")
#install.packages("corrplot")
library(corrplot)
library(psych)
library(tidyverse)
library(skimr)
library(dplyr)
```

## R Markdown

```{r}
EUcars <- read_csv("car_prices.csv", na = c("", NA, "N/A"))

EUcars = EUcars |> drop_na() |> filter (Brand == "Fiat", Gearbox != "Semi-automatic", Drivetrain != "Rear") |> select(Model,Price,Kilometers,Power, Seats,Doors,Cylinders,Gearbox,Year,Drivetrain,Country)

EUcars <- replace(EUcars, EUcars == "Automatic", "0")
EUcars <- replace(EUcars, EUcars == "Manual", "1")
EUcars <- replace(EUcars, EUcars == "4WD", "1")
EUcars <- replace(EUcars, EUcars == "Front", "0")
EUcars <- replace(EUcars, EUcars == "AT", "1")
EUcars <- replace(EUcars, EUcars == "BE", "2")
EUcars <- replace(EUcars, EUcars == "DE", "3")
EUcars <- replace(EUcars, EUcars == "IT", "4")
EUcars <- replace(EUcars, EUcars == "NL", "5")

EUcars$Gearbox = as.numeric(EUcars$Gearbox)
EUcars$Drivetrain = as.numeric(EUcars$Drivetrain)
EUcars$Country = as.numeric(EUcars$Country)

```

```{r}
```

```{r}
# Use eigen() to identify intrinsic dimensionality separtely for predictor and response variables.
predictorMatrix <-EUcars[,c(3,4,6:11)]
responseMatrix <-EUcars[,c(2,5)]

corPredictorMatrix <- cor(predictorMatrix)
corResponseMatrix <- cor(responseMatrix)
```

## Eigen values
```{r}
#Predictor values
eigen(corPredictorMatrix)$values
plot(eigen(corPredictorMatrix)$values)

# For our predictor values using Jolliefs and Kaisers criterion we would get a intrinsic dimensional of 4 while using the inflection point would be 2. Based on this we would use a intrinsic dimensionality of 4.

#Response values
eigen(corResponseMatrix)$values
plot(eigen(corResponseMatrix)$values)

# For our Response variable Kaisers criterion would tell us to use 1 while Jolliefs tells us to use 2 and in this case the inflection point isn't helpful with our graph so based on this we are going to use a intrinsic dimensionality of 1 because that is the only way we can reduce our dimensions for Response.
```



corResponseMatrix <- cor(responseMatrix)
eigen(corResponseMatrix)$values
```
# This criterion is linked to a Learning OutcomeUse pca() to extract the optimally orthogonally and obliquely rotated loading matrices separately for predictor and response variables

## Predictor Variables
```{r}
A <- pca(r = corPredictorMatrix, nfactors = 4, rotate = "varimax")$loadings[]
pca(r = corPredictorMatrix, nfactors = 4, rotate = "varimax")

A3 <- pca(r = corPredictorMatrix, nfactors = 4, rotate = "oblimin")$loadings[]
pca(r = corPredictorMatrix, nfactors = 4, rotate = "oblimin")
```
## Response Variables
```{r}
A <- pca(r = corResponseMatrix, nfactors = 1)$loadings[]
pca(r = corResponseMatrix, nfactors = 1)
```
After using PCA we came to the conclusion that it does not make any sense to cluster the response varibles together because the data loss of 36%is greater than the threshold of 30%. 
